### 2024-03-18 misc/expo/entropy.tex
Ahoy

\proof First we will work in the case where there exists $n$ such that $\pr\{X = a\}$ is a multiple of
$1/n$ for all $a\in A$. Let $Y$ be uniformly distributed on $[n]$ and let $\{E_a\}_{a\in A}$
be a partition of $[n]$ such that $|E_a| = n\pr\{X= a\}$ for all $a\in A$, and let $Z = a$
if $Y\in E_a$. This definition makes $Z$ and $X$ identically distributed, so $\Eta\{Z\} = \Eta\{X\}$
by the invariance axiom, and it suffices to prove $\Eta\{Z\} \ge 0$.

For every $a\in A$, the conditional entropy $\Eta\{Y \given Z = a\}$ is uniformly
distributed on a set of size $|E_a|$. From our choice of $a^*$
we have $|E_{a^*}| \ge |E_a|$ for all $a\in A$. Hence by Proposition~{\proplogeq}, we have
$$\Eta\{Y\given Z\} = \sum_{a\in A} \pr\{X = a\} \Eta\{Y\given X = a\}
\sum_{a\in A} \pr\{X = a\} \log |E_a| \le \log |E_{a^*}|.$$
Since $Z$ is determined by $Y$, we have $\Eta\{Y\} = \Eta\{Z\} + \Eta\{Y\given Z\}$ by the previous lemma.
So by another invocation of Proposition~{\proplogeq}, we have
$$\eqalign{
\Eta\{Z\} &= \Eta\{Y\} - \Eta\{Y\given Z\} \cr
&\ge \log n - \log |E_{a^*}| \cr
&\ge \log \biggl({n\over |E_{a^*}|}\biggr) \cr
&= \log\biggl({1\over \pr\{X = a^*\}}\biggr), \cr
}$$
and hence $2^{-\Etaseven\{X\}} = 2^{-\Etaseven\{Z\}} \le \pr\{X = a^*\}$.

@@

The submodularity inequality is often stated in terms of a triple of random variables in terms of
the {\it conditional mutual information}, which is defined by
$$\eqalign{
\II\{ X : Y \given Z\} &= \sum_{z\in C} p_Z(z) \II\bigl\{ (X\given Z = z) : (Y\given Z = z)\bigr\} \cr
&= \Eta\{X\given Z\} - \Eta\{X\given Y, Z\} \cr
&= \Eta\{Y\given Z\} - \Eta\{Y\given X, Z\}.\cr
}$$

\proclaim Proposition \advthm. Let $X$, $Y$, and $Z$ be discrete random variables. Then
$$\Eta\{X,Y,Z\} + \Eta\{Z\} \le \Eta\{X,Z\} + \Eta\{Y,Z\},$$
which is equivalent to
$$\II\{X : Y\given Z\} \ge 0.$$

@@

To prove the first inequality, note that $(X,Y,Z)$ is jointly determined by $(X,Z)$ and $(Y,Z)$,
and $Z$ is determined by both $(X,Z)$ and $(Y,Z)$ separately, then apply the previous proposition.
Now by the definition of conditional mutual information,
$$\eqalign{
\II\{ X : Y \given Z\} &= \sum_{z\in C} p_Z(z) \II\bigl\{ (X\given Z = z) : (Y\given Z = z)\bigr\} \cr
&= \Eta\{ X\given Z\} + \Eta\{Y\given Z\} - \Eta\{ X, Y \given Z \} \cr
&= \Eta\{ X, Z\} -2\Eta\{ Z\} + \Eta\{Y,Z\} - \Eta\{ X, Y ,Z \} + \Eta\{Z\}\cr
&= \Eta\{ X, Z\} + \Eta\{Y,Z\} - \Eta\{ X, Y ,Z \} - \Eta\{Z\},\cr
}$$
and this proves that the first statement is equivalent to the second.\slug

@@

Now we will examine the case where the random variables in question take values in an abelian group $G$,
meaning we can take sums $X+Y$ and differences $X-Y$ of them. Note that if we condition on $Y$,
then the values taken $X+Y$ are in bijection with values taken by $X$. This leads to the following
proposition.

@@

By submodularity, we have $\II\bigl\{ (X-Y : Z) \given X-Z\bigr\} \ge 0$, so
\edef\eqfirstsubmodularity{\the\eqcount}
$$\eqalign{
0 &\le \II\bigl\{ (X-Y : Z) \given X-Z\bigr\} \cr
&\le \Eta\{X-Y \given X-Z\} + \Eta\{ Z\given X-Z\} - \Eta\{ X-Y, Z \given X-Z\} \cr
&\le \Eta\{X-Y, X-Z\} + \Eta\{ Z, X-Z\} - \Eta\{ X-Y, Z, X-Z\} - \Eta\{X-Z\}.\cr
}\adveq$$
Now, since the values $(x-y,x-z)$ taken by $(X-Y,X-Z)$ are in bijection with values
$(x-z,y-z)$ taken by $(X-Z, Y-Z)$ via the map $(v,w) \mapsto (w, w-v)$,
by the invariance axiom we have
$$\Eta\{X-Y, X-Z\} = \Eta\{X-Z, Y-Z\},$$
and
$$\Eta\{X-Y, X-Z\} \le \Eta\{X-Y\} + \Eta\{Y-Z\}$$
follows by submodularity. Similar invocations of the invariance axiom give
$$\Eta\{Z, X-Z\} = \Eta\{X,Z\}$$
and
$$\Eta\{X-Y, Z, X-Z\} = \Eta\{X,Y,Z\} = \Eta\{X,Z\} + \Eta\{Y\},$$
where in the latter statement the second equality follows from the fact that $(X,Y)$ and $Z$ are
independent.
Substituting these three inequalities into~\refeq{\eqfirstsubmodularity}, we have
$$0\le \Eta\{X-Y\} + \Eta\{Y-Z\} +\Eta\{X,Z\} -\Eta\{X,Z\} + \Eta\{Y\} - \Eta\{X-Z\},$$
whence
$$ \Eta\{X-Z\} \le \Eta\{X-Y\} + \Eta\{Y-Z\} + \Eta\{Y\},$$
which completes the proof.\slug

@@

\proof Since
$$ k = \dd\{X_1, X_2\} = \dd\{X_1, \bar X_2\} = \dd\{X_2, \bar X_1\},$$
applying Lemma~{\lemfivetwo} four times gives us the inequalities
$$\eqalign{
\dd\{X_1^*; X_1 + \bar X_2\}-\dd\{X_1^*, X_1\}&\le{1\over 2}k + {1\over 4}\Eta\{X_2\}-{1\over 4}\Eta\{X_1\},\cr
\dd\{X_2^*; X_2 + \bar X_1\}-\dd\{X_2^*, X_2\}&\le{1\over 2}k + {1\over 4}\Eta\{X_1\}-{1\over 4}\Eta\{X_2\},\cr
\dd\{X_1^*; X_1\given X_1-\bar X_2\}-\dd\{X_1^*,X_1\}
  &\le{1\over 2}k + {1\over 4}\Eta\{X_1\} - {1\over 4}\Eta\{X_2\},\cr
}$$
and
$$\dd\{X_2^*; X_2\given X_2-\bar X_1\}-\dd\{X_2^*,X_2\}
  \le{1\over 2}k + {1\over 4}\Eta\{X_2\} - {1\over 4}\Eta\{X_1\}.$$
By summing these four inequalites, we obtain
\edef\eqfoursum{\the\eqcount}
$$\eqalign{
2k &\ge \dd\{X_1^*; X_1 + \bar X_2\} + \dd\{X_2^*; X_2 + \bar X_1\} \cr
&\qquad\qquad+\dd\{X_1^*; X_1\given X_1 + \bar X_2\} + \dd\{X_2^*; X_2\given X_2 + \bar X_1\} \cr
&\qquad\qquad-2\dd\{X_1^*, X_1\} - 2\dd\{X_2^*, X_2\} .\cr
}\adveq$$
On the other hand,
Corollary~{\corfourtwo} with $(Y_1, Y_2, Y_3, Y_4)$ set to $(X_1, X_2, \bar X_2, \bar X_1)$
gives us
$$\eqalign{
\dd\{X_1, X_2\} + \dd\{\bar X_2,\bar X_1\} = \dd\{&X_1 - \bar X_2;\ X_2 - \bar X_1\} \cr
&+ \dd\{ X_1\given X_1 - \bar X_2;\ X_2\given X_2-\bar X_1\}\cr
&+ \II\{X_1 - X_2 : X_2-\bar X_1 \given X_1 - X_2 - \bar X_2 + \bar X_1\}. \cr
}$$
which can be rewritten
\edef\eqioneidentity{\the\eqcount}
$$\eqalign{
2k &=
\dd\{X_1 + \bar X_2;\ X_2 + \bar X_1\} +\dd\{ X_1\given X_1 + \bar X_2;\ X_2\given X_2+\bar X_1\}\cr
&\qquad\qquad+ \II\{X_1 + X_2 : X_2+\bar X_1 \given X_1 + X_2 + \bar X_1 + \bar X_2\}\cr
&=  \dd\{X_1 + \bar X_2;\ X_2 + \bar X_1\} + \dd\{ X_1\given X_1 + \bar X_2;\ X_2\given X_2+\bar X_1\}+I_1\cr
}\adveq$$
since $\dd\{X_1,X_2\} = k$ and we are working in $G = \FF_2^n$.
By Lemma~{\lemionehelper}, we have
$$\eqalign{
\dd\{X_1 + \bar X_2; X_2 + \bar X_1\} &\ge k-\eta\bigl(\dd\{X_1^*, X_1 + \bar X_2\}-\dd\{X_1^*, X_1\}\cr
&\qquad + \dd\{X_2^*, X_2 + \bar X_1\}- \dd\{X_2^*, X_2\}\bigr) \cr
}\adveq$$
and
\edef\eqioneconditional{\the\eqcount}
$$\eqalign{
\dd\{X_1\given X_1 + \bar X_2;X_2\given X_2 + \bar X_1\}
&\ge k-\eta\bigl(\dd\{X_1^*; X_1\given  X_1 + \bar X_2\} - \dd\{X_1^*, X_1\}\cr
&\qquad+ \dd\{X_2^*, X_2\given X_2 + \bar X_1\}- \dd\{X_2^*, X_2\} \bigr).\cr
}\adveq$$
Substituting these two inequalities into~\refeq{\eqioneidentity} yields
$$\eqalign{
2k &\ge I_1 + 2k - \eta\bigl( \dd\{X_1^*, X_1 + \bar X_2\} + \dd\{X_1^*, X_1\given X_1+\bar X_2\} \cr
&\qquad\qquad\qquad\qquad+ \dd\{X_2^*, X_2 + \bar X_1\} + \dd\{X_2^*, X_2\given X_2+\bar X_2\},\cr
&\qquad\qquad\qquad\qquad -2\dd\{X_1^*, X_1\} - 2\{X_2^*, X_2\} \bigl)\cr
}$$
whence
$$\eqalign{
I_1&\le \eta\bigl( \dd\{X_1^*, X_1 + \bar X_2\} + \dd\{X_1^*, X_1\given X_1+\bar X_2\} \cr
&\qquad\qquad\qquad\qquad+ \dd\{X_2^*, X_2 + \bar X_1\} + \dd\{X_2^*, X_2\given X_2+\bar X_2\} ,\cr
&\qquad\qquad\qquad\qquad -2\dd\{X_1^*, X_1\} - 2\{X_2^*, X_2\} \bigl)\cr
}$$
and we can substitute the inequality~\refeq{\eqfoursum} to get $I_1\le 2\eta k$.

To prove the other claim, we substitute~\refeq{\eqioneconditional} into~\refeq{\eqioneidentity}
to obtain
$$\eqalign{
k&\ge I_1 + \dd\{X_1 + \bar X_2 ; X_2 + \bar X_1\}
  -\eta\bigl(\dd\{X_1^*; X_1\given  X_1 + \bar X_2\} - \dd\{X_1^*, X_1\}\cr
&\hskip150pt+ \dd\{X_2^*, X_2\given X_2 + \bar X_1\}- \dd\{X_2^*, X_2\} \bigr) \cr
&\ge I_1 + \dd\{X_1 + \bar X_2; X_2 + \bar X_1\} -\eta k, \cr
}$$
where in the second line we have used two of the four inequalities begotten by Lemma~{\lemfivetwo}.
So $(1+\eta)k - I_1 \ge \dd\{X_1 + \bar X_2; X_2 + \bar X_1\}$, which we expand to
$$\eqalign{
(1-\eta)k-I_1 &\ge \Eta\{X_1 + X_2 + \bar X_1 + \bar X_2\} - {1\over 2}\Eta\{X_1 + \bar X_2\} - 
{1\over 2} \Eta\{X_2+\bar X_1\} \cr
&=  \Eta\{X_1 + X_2 + \bar X_1 + \bar X_2\}-\dd\{X_1, X_2\} - {1\over 2} \Eta\{X_1\}-{1\over 2}\Eta\{X_2\}\cr
}$$
by independence and the fact that subtraction is addition in $G$. Since
$\dd\{X_1, X_2\}=k$, rearranging completes
the proof.\slug

@@

>>>>

\proof First we will work in the case where there exists $n$ such that $\pr\{X = a\}$ is a multiple of
$1/n$ for all $a\in A$. Let $Y$ be uniformly distributed on $[n]$ and let $\{E_a\}_{a\in A}$
be a partition of $[n]$ such that $|E_a| = n\pr\{X= a\}$ for all $a\in A$, and let $Z = a$
if $Y\in E_a$. This definition makes $Z$ and $X$ identically distributed, so $\Eta\{Z\} = \Eta\{X\}$
by the invariance axiom.

For every $a\in A$, the conditional entropy $\Eta\{Y \given Z = a\}$ is uniformly
distributed on a set of size $|E_a|$. From our choice of $a^*$
we have $|E_{a^*}| \ge |E_a|$ for all $a\in A$. Hence by Proposition~{\proplogeq}, we have
$$\Eta\{Y\given Z\} = \sum_{a\in A} \pr\{X = a\} \Eta\{Y\given X = a\}
= \sum_{a\in A} \pr\{X = a\} \log |E_a| \le \log |E_{a^*}|.$$
Since $Z$ is determined by $Y$, we have $\Eta\{Y\} = \Eta\{Z\} + \Eta\{Y\given Z\}$ by the previous lemma.
So by another invocation of Proposition~{\proplogeq}, we have
$$\eqalign{
\Eta\{Z\} &= \Eta\{Y\} - \Eta\{Y\given Z\} \cr
&\ge \log n - \log |E_{a^*}| \cr
&\ge \log \biggl({n\over |E_{a^*}|}\biggr) \cr
&= \log\biggl({1\over \pr\{X = a^*\}}\biggr), \cr
}$$
and hence $2^{-\Etaseven\{X\}} = 2^{-\Etaseven\{Z\}} \le \pr\{X = a^*\}$.

@@

The submodularity inequality is often stated in terms of a triple of random variables in terms of
the {\it conditional mutual information}, which is defined by
$$\eqalign{
\II\{ X : Y \given Z\} &= \sum_{z\in C} \pr\{Z = z\} \II\bigl\{ (X\given Z = z) : (Y\given Z = z)\bigr\} \cr
&= \Eta\{X\given Z\} - \Eta\{X\given Y, Z\} \cr
&= \Eta\{Y\given Z\} - \Eta\{Y\given X, Z\}.\cr
}$$

\proclaim Proposition \advthm. Let $X$, $Y$, and $Z$ be discrete random variables. Then
$$\Eta\{X,Y,Z\} + \Eta\{Z\} \le \Eta\{X,Z\} + \Eta\{Y,Z\},$$
which is equivalent to
$$\II\{X : Y\given Z\} \ge 0.$$
Equality holds if and only if $X$ and $Y$ are independent conditional on $Z$.

@@

To prove the first inequality, note that $(X,Y,Z)$ is jointly determined by $(X,Z)$ and $(Y,Z)$,
and $Z$ is determined by both $(X,Z)$ and $(Y,Z)$ separately, then apply the previous proposition.
Now by the definition of conditional mutual information,
$$\eqalign{
\II\{ X : Y \given Z\} &= \sum_{z\in C} \pr\{Z=z\} \II\bigl\{ (X\given Z = z) : (Y\given Z = z)\bigr\} \cr
&= \Eta\{ X\given Z\} + \Eta\{Y\given Z\} - \Eta\{ X, Y \given Z \} \cr
&= \Eta\{ X, Z\} -2\Eta\{ Z\} + \Eta\{Y,Z\} - \Eta\{ X, Y ,Z \} + \Eta\{Z\}\cr
&= \Eta\{ X, Z\} + \Eta\{Y,Z\} - \Eta\{ X, Y ,Z \} - \Eta\{Z\},\cr
}$$
and this proves that the first statement is equivalent to the second.\slug

@@

Now we will examine the case where the random variables in question take values in an abelian group $G$,
meaning we can take sums $X+Y$ and differences $X-Y$ of them. Note that if we condition on $Y$,
then the values taken by $X+Y$ are in bijection with values taken by $X$. This leads to the following
proposition.

@@

By submodularity, we have $\II\bigl\{ (X-Y : Z) \given X-Z\bigr\} \ge 0$, so
\edef\eqfirstsubmodularity{\the\eqcount}
$$\eqalign{
0 &\le \II\bigl\{ (X-Y : Z) \given X-Z\bigr\} \cr
&\le \Eta\{X-Y \given X-Z\} + \Eta\{ Z\given X-Z\} - \Eta\{ X-Y, Z \given X-Z\} \cr
&\le \Eta\{X-Y, X-Z\} + \Eta\{ Z, X-Z\} - \Eta\{ X-Y, Z, X-Z\} - \Eta\{X-Z\}.\cr
}\adveq$$
Now, since the values $(x-y,x-z)$ taken by $(X-Y,X-Z)$ are in bijection with values
$(x-z,y-z)$ taken by $(X-Z, Y-Z)$ via the map $(v,w) \mapsto (w, w-v)$,
by the invariance axiom we have
$$\Eta\{X-Y, X-Z\} = \Eta\{X-Z, Y-Z\} \le \Eta\{X-Z\} + \Eta\{Y-Z\}.$$
Similar invocations of the invariance axiom give
$$\Eta\{Z, X-Z\} = \Eta\{X,Z\}$$
and
$$\Eta\{X-Y, Z, X-Z\} = \Eta\{X,Y,Z\} = \Eta\{X,Z\} + \Eta\{Y\},$$
where in the latter statement the second equality follows from the fact that $(X,Y)$ and $Z$ are
independent.
Substituting these three inequalities into~\refeq{\eqfirstsubmodularity}, we have
$$0\le \Eta\{X-Y\} + \Eta\{Y-Z\} +\Eta\{X,Z\} -\Eta\{X,Z\} + \Eta\{Y\} - \Eta\{X-Z\},$$
whence
$$ \Eta\{X-Z\} \le \Eta\{X-Y\} + \Eta\{Y-Z\} + \Eta\{Y\},$$
which completes the proof.\slug

@@

\proof
Since
$$ k = \dd\{X_1, X_2\} = \dd\{X_1, \bar X_2\} = \dd\{X_2, \bar X_1\},$$
applying Lemma~{\lemfivetwo} four times gives us the inequalities
$$\eqalign{
\dd\{X_1^*; X_1 + \bar X_2\}-\dd\{X_1^*, X_1\}&\le{1\over 2}k + {1\over 4}\Eta\{X_2\}-{1\over 4}\Eta\{X_1\},\cr
\dd\{X_2^*; X_2 + \bar X_1\}-\dd\{X_2^*, X_2\}&\le{1\over 2}k + {1\over 4}\Eta\{X_1\}-{1\over 4}\Eta\{X_2\},\cr
\dd\{X_1^*; X_1\given X_1-\bar X_2\}-\dd\{X_1^*,X_1\}
  &\le{1\over 2}k + {1\over 4}\Eta\{X_1\} - {1\over 4}\Eta\{X_2\},\cr
}$$
and
$$\dd\{X_2^*; X_2\given X_2-\bar X_1\}-\dd\{X_2^*,X_2\}
  \le{1\over 2}k + {1\over 4}\Eta\{X_2\} - {1\over 4}\Eta\{X_1\}.$$
By summing these four inequalites, we obtain
\edef\eqionefoursum{\the\eqcount}
$$\eqalign{
2k &\ge \dd\{X_1^*; X_1 + \bar X_2\} + \dd\{X_2^*; X_2 + \bar X_1\} \cr
&\qquad\qquad+\dd\{X_1^*; X_1\given X_1 + \bar X_2\} + \dd\{X_2^*; X_2\given X_2 + \bar X_1\} \cr
&\qquad\qquad-2\dd\{X_1^*, X_1\} - 2\dd\{X_2^*, X_2\} .\cr
}\adveq$$
On the other hand,
Corollary~{\corfourtwo} with $(Y_1, Y_2, Y_3, Y_4)$ set to $(X_1, X_2, \bar X_2, \bar X_1)$
gives us
$$\eqalign{
\dd\{X_1, X_2\} + \dd\{\bar X_2,\bar X_1\} = \dd\{&X_1 - \bar X_2;\ X_2 - \bar X_1\} \cr
&+ \dd\{ X_1\given X_1 - \bar X_2;\ X_2\given X_2-\bar X_1\}\cr
&+ \II\{X_1 - X_2 : X_2-\bar X_1 \given X_1 - X_2 - \bar X_2 + \bar X_1\}. \cr
}$$
which can be rewritten
\edef\eqioneidentity{\the\eqcount}
$$\eqalign{
2k &=
\dd\{X_1 + \bar X_2;\ X_2 + \bar X_1\} +\dd\{ X_1\given X_1 + \bar X_2;\ X_2\given X_2+\bar X_1\}\cr
&\qquad\qquad+ \II\{X_1 + X_2 : X_2+\bar X_1 \given X_1 + X_2 + \bar X_1 + \bar X_2\}\cr
&=  \dd\{X_1 + \bar X_2;\ X_2 + \bar X_1\} + \dd\{ X_1\given X_1 + \bar X_2;\ X_2\given X_2+\bar X_1\}+I_1\cr
}\adveq$$
since $\dd\{X_1,X_2\} = k$ and we are working in $G = \FF_2^n$.
By Lemma~{\lemionehelper}, we have
$$\eqalign{
\dd\{X_1 + \bar X_2; X_2 + \bar X_1\} &\ge k-\eta\bigl(\dd\{X_1^*, X_1 + \bar X_2\}-\dd\{X_1^*, X_1\}\cr
&\qquad + \dd\{X_2^*, X_2 + \bar X_1\}- \dd\{X_2^*, X_2\}\bigr) \cr
}\adveq$$
and
\edef\eqioneconditional{\the\eqcount}
$$\eqalign{
\dd\{X_1\given X_1 + \bar X_2;X_2\given X_2 + \bar X_1\}
&\ge k-\eta\bigl(\dd\{X_1^*; X_1\given  X_1 + \bar X_2\} - \dd\{X_1^*, X_1\}\cr
&\qquad+ \dd\{X_2^*, X_2\given X_2 + \bar X_1\}- \dd\{X_2^*, X_2\} \bigr).\cr
}\adveq$$
Substituting these two inequalities into~\refeq{\eqioneidentity} yields
$$\eqalign{
2k &\ge I_1 + 2k - \eta\bigl( \dd\{X_1^*, X_1 + \bar X_2\} + \dd\{X_1^*, X_1\given X_1+\bar X_2\} \cr
&\qquad\qquad\qquad\qquad+ \dd\{X_2^*, X_2 + \bar X_1\} + \dd\{X_2^*, X_2\given X_2+\bar X_2\},\cr
&\qquad\qquad\qquad\qquad -2\dd\{X_1^*, X_1\} - 2\{X_2^*, X_2\} \bigl)\cr
}$$
whence
$$\eqalign{
I_1&\le \eta\bigl( \dd\{X_1^*, X_1 + \bar X_2\} + \dd\{X_1^*, X_1\given X_1+\bar X_2\} \cr
&\qquad\qquad\qquad\qquad+ \dd\{X_2^*, X_2 + \bar X_1\} + \dd\{X_2^*, X_2\given X_2+\bar X_2\} ,\cr
&\qquad\qquad\qquad\qquad -2\dd\{X_1^*, X_1\} - 2\{X_2^*, X_2\} \bigl)\cr
}$$
and we can substitute the inequality~\refeq{\eqionefoursum} to get $I_1\le 2\eta k$.

To prove the other claim, we substitute~\refeq{\eqioneconditional} into~\refeq{\eqioneidentity}
to obtain
$$\eqalign{
k&\ge I_1 + \dd\{X_1 + \bar X_2 ; X_2 + \bar X_1\}
  -\eta\bigl(\dd\{X_1^*; X_1\given  X_1 + \bar X_2\} - \dd\{X_1^*, X_1\}\cr
&\hskip150pt+ \dd\{X_2^*; X_2\given X_2 + \bar X_1\}- \dd\{X_2^*, X_2\} \bigr) \cr
&\ge I_1 + \dd\{X_1 + \bar X_2; X_2 + \bar X_1\} -\eta k, \cr
}$$
where in the second line we have used two of the four inequalities begotten by Lemma~{\lemfivetwo}.
So $(1+\eta)k - I_1 \ge \dd\{X_1 + \bar X_2; X_2 + \bar X_1\}$, which we expand to
$$\eqalign{
(1-\eta)k-I_1 &\ge \Eta\{X_1 + X_2 + \bar X_1 + \bar X_2\} - {1\over 2}\Eta\{X_1 + \bar X_2\} - 
{1\over 2} \Eta\{X_2+\bar X_1\} \cr
&=  \Eta\{X_1 + X_2 + \bar X_1 + \bar X_2\}-\dd\{X_1, X_2\} - {1\over 2} \Eta\{X_1\}-{1\over 2}\Eta\{X_2\}\cr
}$$
by independence and the fact that subtraction is addition in $G$. Rearranging terms completes
the proof.\slug

@@

\edef\lemitwobound{\the\sectcount.\the\thmcount}
\proclaim Lemma \advthm. Let $G = \FF_2^n$
and suppose $(X_1, X_2)\in \P(G)^2$ is a minimiser of $\tau$ with $\dd\{X_1, X_2\} = k$.
Then letting
$$I_2 = \II\{ X_1 + X_2 : X_1 + \bar X_1 \given X_1 + X_2 + \bar X_1 + \bar X_2\},$$
where
$X_1$ and $\bar X_1$ be copies of $X_1$ and $X_2$ and $\bar X_2$ be copies of $X_2$ such that $X_1$,
$\bar X_1$, $X_2$, and $\bar X_2$ are all independent,
we have
$$I_2 \le 2\eta k + {2\eta(2\eta k- I_1)\over 1-\eta}.$$

\proof
Applying Lemma~{\lemfivetwo} four times gives us the inequalities
\edef\eqitwofourineq{\the\eqcount}
$$\eqalign{
\dd\{X_1^*; X_1 + \bar X_1\}-\dd\{X_1^*, X_1\}&\le{1\over 2}\dd\{X_1, X_1\},\cr
\dd\{X_2^*; X_2 + \bar X_2\}-\dd\{X_2^*, X_2\}&\le{1\over 2}\dd\{X_2, X_2\},\cr
\dd\{X_1^*; X_1\given X_1-\bar X_1\}-\dd\{X_1^*,X_1\} &\le {1\over 2}\dd\{X_1, X_1\},\cr
}\adveq$$
and
$$\dd\{X_2^*; X_2\given X_2-\bar X_2\}-\dd\{X_2^*,X_2\} \le{1\over 2}\dd\{X_2,X_2\}.\adveq$$
These inequalities and Lemma~{\lemionehelper} together give
\edef\eqitwounconditional{\the\eqcount}
$$\eqalign{
\dd\{X_1 + \bar X_1; X_2 + \bar X_2\} &\ge k-\eta\bigl(\dd\{X_1^*, X_1 + \bar X_1\}-\dd\{X_1^*, X_1\}\cr
&\qquad + \dd\{X_2^*, X_2 + \bar X_2\}- \dd\{X_2^*, X_2\}\bigr) \cr
&\ge k-{\eta\over 2}\dd\{X_1, X_1\} -{\eta\over 2} \dd\{X_2, X_2\} \cr
}\adveq$$
and
\edef\eqitwoconditional{\the\eqcount}
$$\eqalign{
\dd\{X_1\given X_1 + \bar X_1;X_2\given X_2 + \bar X_2\}
&\ge k-\eta\bigl(\dd\{X_1^*; X_1\given  X_1 + \bar X_1\} - \dd\{X_1^*, X_1\}\cr
&\qquad+ \dd\{X_2^*, X_2\given X_2 + \bar X_2\}- \dd\{X_2^*, X_2\} \bigr)\cr
&\ge k-{\eta\over 2}\dd\{X_1, X_1\} -{\eta\over 2} \dd\{X_2, X_2\}. \cr
}\adveq$$
Corollary~{\corfourtwo}, with $(Y_1, Y_2, Y_3, Y_4)$ set to $(X_2, X_1, \bar X_2, \bar X_1)$ this time,
yields
\edef\eqitwoidentity{\the\eqcount}
$$\eqalign{
2k &=
\dd\{X_2 + \bar X_2;\ X_1 + \bar X_1\} +\dd\{ X_2\given X_2 + \bar X_2;\ X_1\given X_1+\bar X_1\}\cr
&\qquad\qquad+ \II\{X_1 + X_2 : X_1+\bar X_1 \given X_1 + X_2 + \bar X_1 + \bar X_2\}\cr
&=  \dd\{X_1 + \bar X_1;\ X_2 + \bar X_2\} + \dd\{ X_1\given X_1 + \bar X_1;\ X_2\given X_2+\bar X_2\}+I_2,\cr
}\adveq$$
into which we substitute~\refeq{\eqitwounconditional} and~\refeq{\eqitwoconditional} to obtain
\edef\eqitwoalmostdone{\the\eqcount}
$$ I_2 \le \eta\bigl( \dd\{X_1, X_1\} + \dd\{X_2,X_2\}\bigr).$$
It remains to bound $\dd\{X_1, X_1\} + \dd\{X_2,X_2\}$.

Since $\Eta\{X_1 + \bar X_1\}= \Eta\{X_1\} + \dd\{X_1, X_1\}$ and similarly for $X_2$, we may expand
$$\eqalign{
\dd\{X_1 + \bar X_1; X_2 + \bar X_2\} &= \Eta\{X_1 + \bar X_1 + X_2 + \bar X_2\} \cr
&\qquad\qquad-{1\over 2} \Eta\{X_1+\bar X_1\} - {1\over 2} \Eta\{X_2+\bar X_2\} \cr
&= \Eta\{X_1 + \bar X_1 + X_2 + \bar X_2\} - {1\over 2}\Eta\{X_1\}
-{1\over 2} \Eta\{X_2\} \cr
&\qquad\qquad-{1\over 2} \dd\{X_1,X_1\} - {1\over 2} \Eta\{X_2,X_2\}.\cr
}$$
But recall that the second conclusion of Lemma~{\lemionebound} was
$$\Eta\{X_1 + X_2 + \bar X_1 + \bar X_2\} \le {1\over 2} \Eta\{X_1\} + {1\over 2} \Eta\{X_2\}+(2+\eta)k - I_1,$$
meaning that
$$\dd\{X_1+\bar X_1; X_2+\bar X_2\}\le (2+\eta)k - {1\over 2}\dd\{X_1, X_1\} -{1\over 2}\dd\{X_2,X_2\}-I_1.$$
Chaining this with~\refeq{\eqitwounconditional} yields
$$ k-{\eta\over 2}\dd\{X_1,X_1\} - {\eta\over 2}\dd\{X_2,X_2\} \le
(2+\eta)k-{1\over 2}\dd\{X_1, X_1\} -{1\over 2}\dd\{X_2,X_2\}-I_1,$$
which simplifies to
$$\dd\{X_1,X_1\} + \dd\{X_2,X_2\} \le {2k +2\eta k - 2I_1\over 1-\eta} = 2k + {2(2\eta k -I_1)\over 1-\eta} $$
and hence
$$I_2 \le 2\eta k + {2\eta(2\eta k -I_1)\over 1-\eta}.\noskipslug$$
